\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-08-24 casestudies.tex
some restracturing to emphasize the case studies early on

% cheng to review
\subsection{Distributed Extent Management System}
\label{sec:cases:azurestore}

We used \psharp to test the \emph{distributed extent management} component of the Windows Azure vNext distributed storage system. This component is responsible for managing the partitioned extent metadata and works as follows.

@@

\subsection{Live Azure Table Migration}

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.  A background ``migrator'' job moves all data from the old table to the new table.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/mocked_migration}
\caption{MigratingTable \psharp correctness test environment.}
\label{fig:mockedmigration}
\end{figure}

@@

Since we were designing a new concurrent protocol that we expected to become increasingly complex over time as we add optimizations, we planned from the beginning to maintain a \psharp test harness along with the protocol to maintain confidence in its correctness.

\subsubsection{Modeling approach}
% TBD where this ends up in the paper ~ Matt 2015-08-17

MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.  MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.  Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances with the same backend tables, the behavior complies with the specification of IChainTable2 for the combined input history.

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read) in parallel with the migrator job, there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call completes, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \psharp \term{tables machine} containing all three SpecTables, a collection of \term{service machines} containing identically configured MigratingTables, and a \term{migrator machine} that performs the background migration (Figure~\ref{fig:mockedmigration}).  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  To ensure that the reference table is never observed to be out of sync with the backend tables, after the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

% Draft
We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.  Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.  By default, all event processing occurs on the .NET thread pool.

We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.  Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.

>>>>

\subsection{Azure Storage vNext}
\label{sec:cases:azurestore}

Azure Storage vNext is the next generation storage system for Windows Azure. Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node (see Figure~\ref{fig:azurestore}). Each extent manager is responsible for managing a subset of the extent nodes. Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager. When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/azurestore}
\caption{Top-level components of a distributed extent management system for Windows Azure.}
\label{fig:azurestore}
\end{figure}

@@

The liveness property that must be always eventually satisfied in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent. There was an actual bug in the system, that led this property to fail for some very rare executions. Although this buggy behavior would be observed from time to time during stress testing, there was no way to reproduce the bug. We discuss later in this paper, how we are able to detect and reproduce this bug using our approach.

Azure Storage vNext is very challenging to test. All the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time. However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions. The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied. The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.

\subsection{Live Azure Table Migration}
\label{sec:cases:migration}

The Live Azure Table Migration is a library capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.  MigratingTable provides a \emph{virtual table} with an API similar to that of an ordinary Azure table, backed by a pair of \emph{old} and \emph{new}  tables.  A background \emph{migrator} job moves all data from the old table to the new table.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

@@

@@

\subsection{Azure Service Fabric}
\label{sec:cases:fabric}

Draft.

### 2015-08-24 intro.tex
some restracturing to emphasize the case studies early on

We present two case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol. Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time as they did not have an effective way to reproduce the bug and nail down the culprit. \psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.

@@

\begin{itemize}
\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms.
\item Our infrastructure can test production code written in \csharp, which is a mainstream language.
\item We present two case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
\item ...
\end{itemize}

>>>>

We present three case studies of using \psharp to test production distributed systems for Windows Azure inside Microsoft: a distributed storage management system and a live migration protocol. Using \psharp, we managed to uncover a very subtle bug that was haunting developers for a long time as they did not have an effective way to reproduce the bug and nail down the culprit. \psharp uncovered this bug in a very small setting, which made it easy to examine traces, identify and eventually fix the problem.

@@

\begin{itemize}
\item We present a methodology that allows flexible modeling of the environment of a distributed system using simple language mechanisms.
\item Our infrastructure can test production code written in \csharp, which is a mainstream language.
\item We present three case studies of using \psharp to test production distributed systems, finding bugs that could not be found with traditional testing techniques.
\item ...
\end{itemize}

### 2015-08-24 methodology.tex
some restracturing to emphasize the case studies early on

@@

>>>>

Our goal in this work is to \emph{test what is being executed}. To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework. The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.

\subsection{The \psharp framework}
\label{sec:method:psharp}

\psharp~\cite{deligiannis2015psharp} provides an \emph{event-driven asynchronous programming} language and a \emph{systematic concurrency testing} engine for developing highly-reliable distributed systems.

The \psharp language is an extension of \csharp, built on top of Microsoft's Roslyn\footnote{\url{https://github.com/dotnet/roslyn}} compiler, that enables asynchronous programming using communicating state-machines. \psharp machines can interact asynchronously by sending and receiving events,\footnote{We use the word ``event'' and ``message'' interchangeably.} an approach commonly used to develop distributed systems. This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g. Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).

A \psharp machine consists of an input event queue, states, state transitions, event handlers, fields and methods. Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler. This handler might update a field, create a new machine, or send an event to another machine. In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it. All this functionality is provided in a lightweight runtime library, build on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.

Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code. Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.

A key capability of the \psharp runtime is that it can run in \emph{bug-finding mode}, where a embedded systematic testing engine captures and takes control of all sources of nondeterminism (such as event handler interleavings, failures, and client requests) in a \psharp program, and then systematically explores all possible executions to discover bugs.

\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.

\PDComment{Do we want to actually mention (and compare with) P in this paper?}
%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

\subsection{Methodology}
\label{sec:method:new}

@@

Another important operation that we had to model in order to capture the nondeterminism in  the Azure Storage vNext system, and be able to detect the liveness violation, was to model the timer operation that sends a heartbeat every 5 seconds and a sync message every 5 minutes.

### 2015-08-24 overview.tex
updates in overview
refactoring
some restracturing to emphasize the case studies early on

As an example of a distributed system, Figure~\ref{fig:azurestore} shows the top-level components of Azure Storage vNext, a distributed extent management system for Windows Azure that is used in production inside Microsoft. We will use this system as the main running example in this paper. The system is described in more details in Section~\ref{sec:cases:azurestore}, together with another case study from Microsoft.

Azure Storage vNext consists of multiple extent managers, extent nodes, and network engines that are able to send messages across the network, and enqueue any received messages in the input queue of the corresponding node. Each extent manager is responsible for managing a subset of the extent nodes. Each extent node stores its corresponding extent in a local storage, and sends periodical heartbeats and synchronization messages to the extent manager. When the extent manager receives a synchronization message it is responsible to update extent nodes with the latest extent.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/azurestore}
\caption{Top-level components of a distributed extent management system for Windows Azure.}
\label{fig:azurestore}
\end{figure}

\subsection{Bugs in distributed systems}
\label{sec:overview:bugs}

In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication. To complicate matters further, unexpected failures are the norm in production: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system. Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.

All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques.

@@

\noindent
A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state. As an example, in Azure Storage vNext we assert that whenever a message gets dequeued there must be an action that can handle the received message. Note that this safety property is not specific to Azure Storage vNext, and can be used in any system based on message passing.

Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states. The liveness property that is being checked in the Azure Storage vNext system is that a user-defined $N$ number of extent nodes must be always eventually available with the latest extent.

Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.

\PDComment{Quickly say how we found Cheng's bug?}

\subsection{Testing production systems}
\label{sec:overview:testing}

Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code. However, these techniques are not effective for testing distributed systems, which have many sources of nondeterminism.

As an example, all the unit test and integration test suites of the Azure Storage vNext project successfully pass every single time. However, the developers found that the stress test suite, which constantly kills and launches extent nodes, could fail from time to time after very long executions. The observed failure was that the liveness property described in Section~\ref{sec:overview:bugs} would not get satisfied. The developers had no way to deterministically reproduce this bug or be able to detect what is the culprit, as the traces they were getting were very long and hard to parse.

Our goal in this work instead is to \emph{test what is actually being executed}. To achieve this we developed a new software engineering methodology for testing distributed systems, implemented on top of the \psharp framework. The original \psharp framework is explain in Section~\ref{sec:overview:psharp}, while the new approach is explained in Section~\ref{}.

\subsection{The \psharp approach}
\label{sec:overview:psharp}

The \psharp~\cite{deligiannis2015psharp} framework provides an \emph{event-driven asynchronous programming} language and a \emph{concurrency unit testing} engine for developing highly-reliable distributed systems.

The \psharp language is an extension of \csharp that enables asynchronous programming using communicating state-machines. \psharp machines can interact asynchronously by sending and receiving events, an approach commonly used to develop distributed systems. This programming model is similar to actor-based approaches provided by other asynchronous programming languages (e.g. Scala~\cite{odersky2008programming} and Erlang~\cite{armstrong1996erlang}).

A \psharp machine consists of an input queue for received events, states, state transitions, event handlers, fields and methods. Machines run concurrently with each other, each executing an event handling loop that dequeues an event from the input queue and handles it by invoking an appropriate event handler. This handler might update a field, create a new machine, or send an event to another machine. In \psharp, a send operation is non-blocking; the message is simply enqueued into the input queue of the target machine, and it is up to the operating system scheduler to decide when to dequeue an event and handle it.

Because \psharp is built on top of \csharp, the programmer can blend \psharp and \csharp code; this not only lowers the overhead of learning a new language, but also allows \psharp to easily integrate with legacy code. Another advantage is that the programmer can use the familiar programming and debugging environment of Visual Studio.

The second core component of the \psharp project is a concurrency unit testing engine that is embedded in the \psharp runtime. This engine allows the event schedulings of a \psharp program to be explored in a controlled manner, facilitating deterministic replay of bugs.

\psharp is available as open-source\footnote{\url{https://github.com/p-org/PSharp}} and is currently used by various teams in Microsoft to develop and test distributed protocols and systems.

%The \psharp language belongs to the same family of languages as P~\cite{desai2013p}.

>>>>

\subsection{Challenges in testing}
\label{sec:bg:challenges}

In a distributed system, message handlers can interleave in arbitrary order, because of the asynchronous nature of message-based communication. To complicate matters further, unexpected failures are the norm in production systems: nodes in a cluster might fail at any moment, and thus programmers have to implement sophisticated mechanisms that can deal with these failures and recover the state of the system. Moreover, with multicore machines having become a commodity, individual components of a distributed system are commonly implemented using multithreaded code, which adds another source of nondeterminism.

All the above sources of nondeterminism (as well as nondeterminism due to timeouts, message losses and client requests) can easily create \emph{heisenbugs}~\cite{gray1986computers, musuvathi2008finding}, which are corner-case bugs that are difficult to detect, diagnose and fix, without using advanced \emph{asynchrony-aware} testing techniques. Techniques such as unit testing, integration testing and stress testing are heavily used in industry today for finding bugs in production code. However, these techniques are not effective for testing distributed systems, as they are not able to capture and control the many sources of nondeterminism.

\subsection{Types of bugs}
\label{sec:bg:bugs}

@@

\noindent
A safety property can be specified using an \emph{assertion} that fails if the property gets violated in some program state. An example of a generic safety property for message passing systems is to assert that whenever a message gets dequeued there must be an action that can handle the received message.

Liveness properties are much harder to specify and check since they apply over entire program executions and not just individual program states. Normally, liveness checking requires the identification of an infinite fair execution that never satisfies the liveness property~\cite{schuppan2004efficient, musuvathi2008fair}. Prior work~\cite{schuppan2004efficient} has proposed that assuming a program with finite state space, a liveness property can be converted into a safety property. Other researchers proposed the use of heuristics and only exploring finite executions of an infinite state space system using random walks to identify if a liveness property is violated~\cite{killian2007life}.
